### 2016-03-30 ch05-initvals/initial-values.tex
Some more uniformity in figure + text

\item \emph{Quadratic approximations of the ML}

@@

A reliably good estimate of the FRF of the chosen system can be obtained via the LPM, which is summarized as follows~\citep{Lumori2014TIM}. Formulated as a nonparametric linear-least-squares-estimate, the LPM is first applied to estimate the FRF from a full data record of a SISO system, systematically expressed in an output-error framework, shown in \figref{fig:oesetup}. The smooth characteristics of both the exact \gls{FRF} $\true{G}$ and the transient term $T$ allow for an optimal application of the \gls{LPM}, leading to a smooth \gls{FRF} estimate with the transient suppressed.

@@

%The estimation of $\hat{G}_\text{poly}(\Omega_k)$ at a single frequency $\Omega_k = 0.586$~Hz is illustrated in Fig. \ref{LPM_Schematic_EG}.

\emph{Impulse response truncation}. 
%\comment{JL: in this paper, the impulse response truncation as an initial estimator is considered together with the LPM. So I would not start a new subsection for it.}
The estimate is smoothed further by truncating its impulse response function $g_{\text{poly}}(t) = \mathrm{IDFT}\, \left( \hat{G}_\text{poly}(\Omega_k)) \right)$ as originally presented in~\citep{Lumori2014TIM}.
%\comment{JL: $\hat{G}_\text{poly}(\Omega_k))$ is undefined}.

The impulse response is truncated after the time $\tau_{\mathrm{trunc}}$ where the signal becomes indistinguishable from the noise, i.e.,
\begin{equation}
  g_{\mathrm{trunc}}(t) = 
  \begin{cases}
    g_{\text{poly}}(t) & t \leq \tau_{\mathrm{trunc}} \\
    0                    & t \geq \tau_{\mathrm{trunc}}
  \end{cases}
  \text{.}
\end{equation}
To this end, an estimate of the envelope of impulse response is determined by fitting an exponential function $g_{\mathrm{exp}}(t) = B e^{\beta t}$ to the peaks of the impulse response $\abs{g_{\mathrm{poly}}(t)}$ using a linear least-squares approach.
Then, $\tau_{\mathrm{trunc}}$ is determined as the time instant where this envelope function sinks below the noise level $\sigma_g$ of the impulse response, i.e. 
\begin{equation}
  \tau_{\mathrm{trunc}} = \min \left\{  t :  g_{\mathrm{exp}}(t) < \gamma \sigma_g \right\}
  = \beta^{-1} \ln \frac{\gamma \sigma_g}{B}
  \text{,}
\end{equation}
where $\gamma=1$ was used in this chapter for simplicity.
By changing $\gamma$, the user can fine-tune the bias/variance trade-off of the estimated FRF $G_{\mathrm{trunc}}(\Omega_k) = \mathrm{DFT} \left( g_{\mathrm{trunc}}(t)\right)$ further.
This can lead to a significant improvement over the classical LPM~\citep{Lumori2014TIM}.

@@

%\JL{
The \gls{RFIR} estimator is formulated in the time domain. 
It estimates the impulse response of a discrete time system as the minimizer of the following regularized least squares objective function:
\begin{align}\label{eq:hRFIRdef}
\hat g_{\mathrm{RFIR}} &= \argmin_{g} \norm{y - g \ast u }^2 + \sigma^2 g^{\TT}P^{-1}g
\end{align}
where $g$ is the vectorized impulse response $g(t)$, with $t = 0,1,\cdots,n_h - 1$, assuming that the impulse response is $n_g$ samples long. 
Furthermore, $g\ast u$ is the convolution of $g$ with the input signal $u$, and $P\in \mathbb{R}^{n_g\times n_g}$ is the kernel matrix~\citep{Chen2012}. 
Note that \eqref{eq:hRFIRdef} is quadratic in $g$, such that $\hat g_{\mathrm{RFIR}}$ can be computed analytically in a single step.
The kernel matrix $P$ embodies prior knowledge on the system to be estimated. 
Here, the \gls{DCkernel}~\citep{Chen2012,Pillonetto2010} is used. 
Specifically, the element at $t_1,t_2$ of $P$ is given by
\begin{equation}
P(\alpha,\beta)_{t_1,t_2} = e^{-\alpha|t_1 - t_2|}e^{-\beta(t_1 + t_2)/2}
\end{equation}
 for $t_1,t_2 \in \left\{0,1,\cdots,n_h - 1\right\}$ which makes the estimated impulse response to be an exponentially decaying function of $t$ with decay rate $\beta^{-1}$, and correlation length $\alpha^{-1}$. 
 The latter \emph{hyperparameters} $\alpha$, $\beta$ and $\sigma$ are determined using Empirical Bayes~\citep{Carlin2000,Gelman2014}, i.e. as the maximizers of the \gls{LML} of the measured output signal~\citep{Chen2012}:
\begin{align}
\mathrm{LML}(y) = -y^{\TT}\Sigma(\alpha,\beta,\sigma)^{-1}y - \log |\Sigma(\alpha,\beta,\sigma)| \text{,}
\end{align}
where $\Sigma(\alpha,\beta, \sigma) = \sigma^2 \I + \phi^{\TT}P(\alpha,\beta)\phi$ and $\phi$ is a Toeplitz matrix constructed with $u$. 
This problem is solved using \mcode{fmincon} in \MATLAB.

>>>>

\item \emph{Quadratic approximations of the \gls{MLE}}

@@

A good estimate of the \gls{FRF} of the chosen system can be obtained via the truncated \gls{LPM}, which is summarized as follows~\citep{Lumori2014TIM} and \secref{sec:nonparametric:truncation} of this thesis.
Here, we repeat the main points briefly.
The \gls{LPM} is first applied to estimate the \gls{FRF} from a full input/output data record of the \gls{SISO} system.
The smooth characteristics of both the exact \gls{FRF} $\true{G}$ and the transient term $T$ allow for application of the \gls{LPM}, leading to a smooth \gls{FRF} estimate with the transient and noise suppressed.

@@

\emph{Impulse response truncation}. 
The estimate is smoothed further by truncating its impulse response function $g_{\text{poly}}(t) = \mathrm{IDFT}\, \left( \hat{G}_\text{poly}(\Omega_k)) \right)$ as originally presented in~\citep{Lumori2014TIM} and \secref{sec:nonparametric:truncation}.

The impulse response is truncated after the time $\truncTime$ where the signal becomes indistinguishable from the noise, i.e.,
\begin{equation}
  g_{\mathrm{trunc}}(t) = 
  \begin{cases}
    g_{\text{poly}}(t) & t \leq \truncTime \\
    0                              & t \geq \truncTime
  \end{cases}
  \text{.}
\end{equation}

To this end, an estimate of the envelope of impulse response is determined by fitting an exponential function $g_{\mathrm{exp}}(t) = B e^{\beta t}$ to the peaks of $\abs{g_{\mathrm{poly}}(t)}$ using a linear least-squares approach.
Then, $\truncTime$ is determined as the time instant where this envelope function sinks below the noise level $\sigma_g$ of the impulse response, i.e. 
\begin{equation}
  \truncTime = \min \Set{ t :  g_{\mathrm{exp}}(t) < \gamma \sigma_g  }
  = \beta^{-1} \ln \frac{\gamma \sigma_g}{B}
  \text{,}
\end{equation}
where $\gamma=1$ was used in this chapter for simplicity.
By changing $\gamma$, the user can fine-tune the bias/variance trade-off of the estimated \gls{FRF} $G_{\mathrm{trunc}}(\Omega_k) = \mathrm{DFT} \left( g_{\mathrm{trunc}}(t)\right)$ further.
This can lead to a significant improvement over the classical LPM~\citep{Lumori2014TIM}.

@@

The \gls{RFIR} estimator is formulated in the time domain. 
It estimates the impulse response of a discrete time system as the minimizer of the following regularized least squares objective function:
\begin{align}\label{eq:hRFIRdef}
\hat g_{\mathrm{RFIR}} &= \argmin_{g} \norm{y - g \ast u }^2 + \sigma^2 g^{\TT}P^{-1}g
\end{align}
where $g$ is the vectorized impulse response $g(t)$, with $t = 0,1,\cdots,n_h - 1$, assuming that the impulse response is $n_g$ samples long. 
Furthermore, $g\ast u$ is the convolution of $g$ with the input signal $u$, and $P\in \mathbb{R}^{n_g\times n_g}$ is the kernel matrix~\citep{Chen2012}. 
Note that \eqref{eq:hRFIRdef} is quadratic in $g$, such that $\hat g_{\mathrm{RFIR}}$ can be computed analytically in a single step.
The kernel matrix $P$ embodies prior knowledge on the system to be estimated. 
Here, the \gls{DCkernel}~\citep{Chen2012,Pillonetto2010} is used. 
Specifically, the element at $t_1,t_2$ of $P$ is given by
\begin{equation}
P(\alpha,\beta)_{t_1,t_2} = e^{-\alpha|t_1 - t_2|}e^{-\beta(t_1 + t_2)/2}
\end{equation}
 for $t_1,t_2 \in \left\{0,1,\cdots,n_h - 1\right\}$ which makes the estimated impulse response to be an exponentially decaying function of $t$ with decay rate $\beta^{-1}$, and correlation length $\alpha^{-1}$. 
 The latter \emph{hyperparameters} $\alpha$, $\beta$ and $\sigma$ are determined using Empirical Bayes~\citep{Carlin2000,Gelman2014}, i.e. as the maximizers of the \gls{LML} of the measured output signal~\citep{Chen2012}:
\begin{align}
\mathrm{LML}(y) = -y^{\TT}\Sigma(\alpha,\beta,\sigma)^{-1}y - \log |\Sigma(\alpha,\beta,\sigma)| \text{,}
\end{align}
where $\Sigma(\alpha,\beta, \sigma) = \sigma^2 \I + \phi^{\TT}P(\alpha,\beta)\phi$ and $\phi$ is a Toeplitz matrix constructed with $u$. 
This problem is solved using \mcode{fmincon} in \MATLAB.

### 2016-03-30 config/styles.inc.tex
Some more uniformity in figure + text

%% Truncation
\pgfplotsset{truncationline/.style={dashdotdotted,line width=1.5pt,line join=round,line cap=round}}

\colorlet{FSmooth}{TangoScarletRed3}
\colorlet{FSmoothLight}{TangoScarletRed2}
\colorlet{expFit}{TangoSkyBlue3}
\colorlet{expFitLight}{TangoSkyBlue2}
\colorlet{trunc}{clogms}

@@

% adapted from xbar legend from pgfplots manual
\pgfplotsset{logHistogramLegend/.style={/pgfplots/legend image code/.code={%
            \draw[##1,/tikz/.cd,bar width=3pt,yshift=-0.2em,bar shift=0pt]
            plot coordinates {(0.15cm,0.6em) (0.3cm,0.8em) (0.45cm,0.6em)};}
}}

>>>>

%% Truncation
\colorlet{FSmooth}{TangoScarletRed3}
\colorlet{FSmoothLight}{TangoScarletRed2}
\colorlet{expFit}{TangoSkyBlue3}
\colorlet{expFitLight}{TangoSkyBlue2}
\colorlet{trunc}{clogms}

@@

% adapted from xbar legend from pgfplots manual
\pgfplotsset{logHistogramLegend/.style={/pgfplots/legend image code/.code={%
            \draw[##1,/tikz/.cd,bar width=3pt,yshift=-0.2em,bar shift=0pt]
            plot coordinates {(0.15cm,0.6em) (0.3cm,0.8em) (0.45cm,0.6em)};}
}}
% adapted from line legend
\pgfplotsset{verticalLineLegend/.style={
  legend image code/.code={
    \draw[mark repeat=2,mark phase=2,##1]
      plot coordinates {
         (0.3cm, 0cm)
         (0.3cm, 0.25em)
         (0.3cm, 0.5em)
      };%
  }
}}

\colorlet{impulseResponseFull}{TangoScarletRed1!50!white}
\colorlet{impulseResponseAbs}{TangoChameleon1}
\colorlet{impulseResponseTruncated}{TangoChameleon3}
\colorlet{impulseResponsePeaks}{TangoScarletRed3}

\pgfplotsset{imprespFull/.style={color=impulseResponseFull,solid}}
\pgfplotsset{imprespAbs/.style={color=impulseResponseAbs,solid}}
\pgfplotsset{imprespExpFit/.style={color=expFit, solid, line width=1.5pt}}
\pgfplotsset{imprespPeaks/.style={color=TangoScarletRed3,mark=*,only marks,line width=1.25pt,mark size=0.5pt}}
\pgfplotsset{imprespTrunc/.style={color=impulseResponseTruncated,solid}}

\pgfplotsset{imprespDetectLevel/.style={color=black,dashed}}
\pgfplotsset{imprespNoiseLevel/.style={color=black,dashed}}

\pgfplotsset{truncationline/.style={dashdotdotted,line width=1.5pt,line join=round,line cap=round, verticalLineLegend}}
